\titletext{}

### 2021-01-25 manuscript/regional_befr_manuscript.tex
Added TRY species matching

% African savannas are understudied in this regard
The representation of dry tropical ecosystems in the BEFR literature is poor compared to other ecosystems. \citet{Clarke2017} conducted a meta-analysis of 182 published BEFR studies, finding that only 13\% were conducted in the tropics generally, with 42\% of those being conducted in the wet tropical forests of Costa Rica, a narrow geographic region \citep{Barthlott2005}. A severe lack of study in dry tropical ecosystems, especially given the potential divergence in BEFR mechanisms described above, suggests that a focus on the BEFR in southern African woodlands could greatly strengthen our understanding of a global BEFR and its environmental determinants. A small number of studies in southern African woodlands, all of which were restricted in spatial scope to a small region of miombo woodland, found that above-ground woody carbon/biomass stocks correlate positively with tree species richness \citep{McNicol2018, Shirima2015, Mutowo2012}. The results of these fine-scale studies concur with similar studies in other biomes \citep{Cardinale2009}. Studies of the BEFR often find that at fine spatial scales (<1 ha), biodiversity shows a strong effect on ecosystem function, but at broad spatial scales  (>10000s ha) biodiversity effects pale in significance compared to abiotic factors such as climate \citep{Pasari2013}.

@@

% Plots chosen
Plots were chosen from a larger pool of 5395 plots held in the SEOSAW database \citep{seosaw_web} based on the quality and completeness of data collection, and plot setup. Plot vegetation was identified under the broad term of `savannaj', which includes `woodland', `savanna woodland', and `tree savanna', variously defined in other areas of the scientific literature and here referred to collectively as southern African woodlands \citep{Ratnam2011, Hill2010}. Plots with evidence of farming, human resource extraction or experimental treatments such as prescribed burning or herbivore exclusion were excluded from the initial pool. Only plots >0.1 hectares were used in analyses, as area-based biomass estimation from small plots is highly influenced by rare large trees \citep{Stegen2011}, leading to inaccurate estimates. Only plots with a stem density >50 trees ha\textsuperscript{-1} (>10 cm stem diameter) were used, to ensure all plots represented woodland rather than `grassy savanna', which is considered a separate biome with very different species composition \citep{Parr2014}.

@@

In our sub-sampling of the plot dataset by stem density, we found an increasing positive effect of tree species diversity on AGB as stem density increased (\autoref{sem_struc_stems_ha}). There appears to be a minimum stem density threshold at c. 180 trees ha\textsuperscript{-1} below which there appears to be a reasonably constant baseline effect of tree diversity on biomass. The effect of structural diversity on AGB appears to remain constant with increasing stem density. The indirect effect of tree species diversity on AGB via structural diversity climbs as stem density increases.

@@

A model incorporating the latent variables of water availability, soil fertility and disturbance by fire showed that the total effect of tree species diversity on biomass was similar to that of water availability, soil fertility and disturbance (\autoref{full_mod}, Fig. S4). The direct effects of water availability, soil fertility and disturbance on AGB were negligible (water: \fmbetamb{}, soil: \fmbetasb{}, disturbance: \fmbetafb{}), with nearly all of their observed effect on AGB coming from the indirect paths via stem density (water: \fmbetamib{}, soil: \fmbetasib{}, disturbance: \fmbetafib{}) and species diversity (water: \fmbetamd{}, soil: \fmbetasd{}, disturbance: \fmbetafd{}). MAP and soil sand content had the greatest contributions to the latent variable of water availability. Model fit was acceptable: CFI = 0.925, TLI = 0.900, and RMSEA = \fmrmsea{}, R\textsuperscript{2} of AGB = \fmrsq{}.

@@

% 
We found a consistent positive effect of tree species diversity on AGB across all models in the current study. Within southern African woodlands we therefore find support that higher tree species richness and evenness leads to higher woody AGB. This finding is in agreement with many other studies across different ecosystems and biomes, supporting the idea that there is a generalisable positive association between biodiversity and ecosystem function \citep{Liang2016, Cardinale2009}. Our study provides a novel dissection of the mechanisms underlying this relationship, particularly in the context of southern African woodlands, a disturbance-structured and poorly studied ecological system.

@@

@@

\begin{figure}[H]
\centering
	\includegraphics[width=0.8\textwidth]{sem_struc_stems_ha}
	\caption{Line plots showing the variation in SEM path coefficients across datasets with different mean stem density. Smoothed lines are loess curves with $\pm1$ standard error shaded bars.}
	\label{sem_struc_stems_ha}
\end{figure}
\newpage{}

>>>>

% African savannas are understudied in this regard
The representation of dry tropical ecosystems in the BEFR literature is poor compared to other ecosystems. \citet{Clarke2017} conducted a meta-analysis of 182 published BEFR studies, finding that only 13\% were conducted in the tropics generally, with 42\% of those being conducted in the wet tropical forests of Costa Rica, a narrow geographic region \citep{Barthlott2005}. A severe lack of studies in dry tropical ecosystems, especially given the potential divergence in BEFR mechanisms described above, suggests that a focus on the BEFR in southern African woodlands could greatly strengthen our understanding of a global BEFR and its environmental determinants. A small number of studies in southern African woodlands, all of which were restricted in spatial scope to a small region of miombo woodland, found that above-ground woody carbon/biomass stocks correlate positively with tree species richness \citep{McNicol2018, Shirima2015, Mutowo2012}. The results of these fine-scale studies concur with similar studies in other biomes \citep{Cardinale2009}. Studies of the BEFR often find that at fine spatial scales (<1 ha), biodiversity shows a strong effect on ecosystem function, but at broad spatial scales  (>10000s ha) biodiversity effects pale in significance compared to abiotic factors such as climate \citep{Pasari2013}.

@@

% Plots chosen
Plots were chosen from a larger pool of 5395 plots held in the SEOSAW database \citep{seosaw_web} based on the quality and completeness of data collection, and plot setup. Plot vegetation was identified under the broad term of `savannah', which includes `woodland', `savanna woodland', and `tree savanna', variously defined in other areas of the scientific literature and here referred to collectively as southern African woodlands \citep{Ratnam2011, Hill2010}. Plots with evidence of farming, human resource extraction or experimental treatments such as prescribed burning or herbivore exclusion were excluded from the initial pool. Only plots >0.1 hectares were used in analyses, as area-based biomass estimation from small plots is highly influenced by rare large trees \citep{Stegen2011}, leading to inaccurate estimates. Only plots with a stem density >50 trees ha\textsuperscript{-1} (>10 cm stem diameter) were used, to ensure all plots represented woodland rather than `grassy savanna', which is considered a separate biome with very different species composition \citep{Parr2014}.

@@

In our sub-sampling of the plot dataset by stem density, we found an increasing positive effect of tree species diversity on AGB as stem density increased (\autoref{sem_struc_stems_ha}). There appears to be a minimum stem density threshold at c. 180 trees ha\textsuperscript{-1} below which there appears to be a reasonably constant baseline effect of tree diversity on biomass. The effect of structural diversity on AGB appears to remain constant with increasing stem density. The indirect effect of tree species diversity on AGB via structural diversity increases as stem density increases.

@@

A model incorporating the latent variables of water availability, soil fertility and disturbance by fire showed that the total effect of tree species diversity on biomass was similar to that of water availability, soil fertility and disturbance (\autoref{full_mod}, Fig. S4). The direct effects of water availability, soil fertility and disturbance on AGB were negligible (water: \fmbetamb{}, soil: \fmbetasb{}, disturbance: \fmbetafb{}), with nearly all of their observed effects on AGB coming from the indirect paths via stem density (water: \fmbetamib{}, soil: \fmbetasib{}, disturbance: \fmbetafib{}) and species diversity (water: \fmbetamd{}, soil: \fmbetasd{}, disturbance: \fmbetafd{}). MAP and soil sand content had the greatest contributions to the latent variable of water availability. Model fit was acceptable: CFI = 0.925, TLI = 0.900, and RMSEA = \fmrmsea{}, R\textsuperscript{2} of AGB = \fmrsq{}.

@@

% 
We found a consistent positive effect of tree species diversity on AGB. Within southern African woodlands we therefore find support for the hypothesis that higher tree species richness and evenness leads to higher above-ground woody biomass. This finding is in agreement with many other studies across different ecosystems and biomes, supporting the idea that there is a generalisable positive association between biodiversity and ecosystem function \citep{Liang2016, Cardinale2009}. Our study provides a novel dissection of the mechanisms underlying this relationship, particularly in the context of southern African woodlands, a disturbance-structured and poorly studied ecological system.

@@

\section{Conclusions}

@@

\begin{figure}[H]
\centering
	\includegraphics[width=0.8\textwidth]{sem_struc_stems_ha}
	\caption{Line plots showing the variation in SEM path coefficients across datasets with different mean stem density. Smoothed lines are loess curves with $\pm1$ standard error shaded bars. AGB = Above-Ground woody Biomass, arrows in plot titles indicate causal paths in SEM models.}
	\label{sem_struc_stems_ha}
\end{figure}
\newpage{}
